### 2020-08-04 conclusions.tex
operators too

In this paper, we showed that program mutants can be used as a proxy
for real faults, to compare (and motivate improvements to) static
analysis tools.  Mutants are attractive in that changing a mostly
correct program usually introduces a bug into it; this is the basis of
mutation testing, after all, and a large body of work supports the
claim that at least 60-70\% of mutants are fault-inducing.   This
means we can assume mutants are faulty, and escape the
ground-truth/false positive problem that makes comparing static
analysis tools so labor-intensive.  Our approach cannot identify precision
problems directly, but combined with finding counts for un-mutated
code, can compute a mutant ratio that implicates imprecise tools.  We
evaluated 9 popular static analysis tools, for Solidity smart contracts, Java,
and Python, and offer advice to users of these tools.  Our mutation
results strongly confirm the wisdom of using multiple tools; with the
exception of one Java tool, all tools we investigated uniquely
detected over 1,000 mutants, and for Java and Python there were no
mutants detected by all tools. For Solidity,
academic research evaluations of the tools generally agreed strongly
with our conclusions, but lacked the detail mutant analysis contributed.
We were also able to use our methods, plus a novel mutant prioritization scheme, to
identify and implement three useful new detectors for the open source Slither smart contract
analyzer, the best-performing of the tools.

>>>>

In this paper, we showed that program mutants can be used as a proxy
for real faults, to compare (and motivate improvements to) static
analysis tools.  Mutants are attractive in that a large body of work supports the
claim that at least 60-70\% of mutants are fault-inducing.   This
allows us to assume \emph{detected} mutants are faulty, and escape the
ground-truth/false positive problem that makes comparing static
analysis tools so labor-intensive.   We
evaluated 9 popular static analysis tools, for Solidity smart contracts, Java,
and Python, and offer advice to users of these tools.  Our mutation
results strongly confirm the wisdom of using multiple tools; with the
exception of one Java tool, all tools we investigated uniquely
detected over 1,000 mutants, and for Java and Python there were no
mutants detected by all tools. For Solidity,
academic research evaluations of the tools generally agreed strongly
with our conclusions, but lacked the detail mutant analysis contributed.
We were also able to use our methods, plus a novel mutant prioritization scheme, to
identify and implement three useful new detectors for the open source Slither smart contract
analyzer, the best-performing of the tools.

### 2020-08-04 experiments.tex
operators too

Our primary experimental results are a set of comparisons of tools using our method, for three languages: Solidity (the most popular language for smart contracts), Java, and Python.  We used the Universal Mutator tool for all experiments; for Solidity and Python, we believe the Universal Mutator is simply the best available tool.  For Java, PIT \cite{pitmut} is more popular, but does not produce source-level mutants, needed for PMD and for manual inspection of results.
We used our results to answer a set of research questions:

@@

Because examining all mutants manually is infeasible,  for {\bf RQ5} we focused primarily on examining mutants in the \emph{difference}: mutants detected by at least one tool, but not detected by all tools, the only ones that actually influence the comparison of tools.  The vast majority of the mutants in this set were meaningful semantic changes a static analysis tool could be expected to detect, and the findings produced by tools were relevant to the nature of the fault.  We do not believe that all mutants represent definite faults; some are harmless but unusual code changes.  A large number of the instances where use of {\tt tx.origin} in place of {\tt msg.sender} was flagged seem to us to be strange, but not necessarily incorrect, code.  On the other hand, at the risk of false positives, it is not unreasonable for tools to report such strange code.   Our estimate is that, ignoring {\tt tx.origin} cases, \emph{at least 70\% of the mutants detected by one, but not all, tools, represent realistic bugs}, and failure to detect is roughly equally due to missing detectors and imprecise analysis.

@@

For Java, again, the large majority of mutants we inspected for {\bf RQ5}, where at least one tool detected the mutant and at least one tool did not detect it were definitely meaningful.  In particular, for Java, the large majority of mutants involved either deleted method calls or changes to conditionals (e.g., {\tt == null} to {\tt != null}) that would clearly introduce potential null pointer exceptions (NPEs), and such a possible NPE was the produced finding.  While the differences between Solidity tools were often due to different detectors, the Java differences seemed mostly rooted in analysis engine methods; all tools aim to warn about potential NPEs.  Because the number of mutants we could examine and understand was smaller, we are less confident in making a probabilistic estimate than with Solidity, but it was clear the basis of the comparison was primarily realistic faults, which some tools detected and others did not.

@@

For Python, we analyzed the top 25 GitHub projects by our criteria (see above), due to the smaller size of Python projects.  These ranged in size from 3,377 LOC to 34,671 LOC, with a total size of just under 100 KLOC.  For Python, we were able to produce 167,511 valid mutants to analyze.

@@

For Python, as with Java, many mutants we inspected for {\bf RQ5} involved deletion of a call or change in comparison operator, leading to a likely method call or access to a variable containing {\tt None}, and the generated finding reported this problem.  Another example of a possibly engine-based mutation difference is that pyflakes did a much better job than the other tools of detecting cases where a mutant caused an exception to be unwisely ignored; these accounted for more than 10\% of differentially-relevant mutants.  We assume the issue is engine-based in that while pyflakes did by far the best at this (real) problem, there were instances where another tool found the problem and it did not.  On the other hand, pyflakes, in the sample we examined, did worst at detecting mutants that removed a needed argument from a function call, another common pattern (Pylint performed best for this type of mutant).  Again, based on a qualitative examination of many mutants, we believe the results to primarily rest on realistic faults, which some tools correctly detected and others did not.

>>>>

Our primary experimental results are a set of comparisons of tools using our method, for three languages: Solidity (the most popular language for smart contracts), Java, and Python.  We used the Universal Mutator tool for all experiments; for Solidity and Python, we believe the Universal Mutator is simply the best available tool.  For Java, PIT \cite{pitmut} is more popular, but does not produce source-level mutants, needed for PMD and for manual inspection of results.   Universal Mutator includes a large set of mutation operators, some unconventional (e.g., swapping order of function arguments) but based on real-world bugs; the complete set is described by regular expressions at \url{https://github.com/agroce/universalmutator/tree/master/universalmutator/static}.  However, the large majority of generated compiling mutants that were detected came from a small set of  commonly-used operators \cite{mutant}, particularly 1) statement/code deletion and 2) operator, conditional, and constant replacements.

We used our results to answer a set of research questions:

@@

Following the method proposed in Section \ref{sec:method},  for {\bf RQ5} we focused on examining mutants detected by at least one tool, but not detected by all tools, the only ones that actually influence the comparison of tools.  The vast majority of the mutants in this set were meaningful semantic changes a static analysis tool could be expected to detect, and the findings produced by tools were relevant to the nature of the fault.  We do not believe that \emph{all} mutants represent definite faults; some are harmless but unusual code changes.  Many cases where use of {\tt tx.origin} in place of {\tt msg.sender} was flagged seem to us to be strange, but not necessarily incorrect, code.  On the other hand, it is not at all unreasonable for tools to report such notably strange code.   Our estimate is that, ignoring {\tt tx.origin} cases, \emph{at least 70\% of the mutants detected by one, but not all, tools, represent realistic bugs}, and failure to detect is roughly equally due to missing detectors and imprecise analysis.

@@

For Java, again, the large majority ($>$ 75\%) of randomly chosen mutants in the tool difference set we inspected for {\bf RQ5} were definitely meaningful, essentially ``real faults.''  In particular, for Java, the large majority of mutants involved either deleted method calls or changes to conditionals (e.g., {\tt == null} to {\tt != null}) that would clearly introduce potential null pointer exceptions (NPEs), and such a possible NPE was the produced finding.  While the differences between Solidity tools were often due to different detectors, the Java differences seemed mostly rooted in analysis engine methods; all tools aim to warn about potential NPEs.  Because the number of mutants we could examine and understand was smaller, we are less confident in making a probabilistic estimate than with Solidity, but it was clear the basis of the comparison was primarily realistic faults, which some tools detected and others did not.

@@

For Python, we analyzed the top 25 GitHub projects by our criteria (see above), due to the smaller size of Python projects.  These ranged in size from 3,377 LOC to 34,671 LOC, with a total size of just under 100 KLOC.  For Python, we were able to produce 167,511 valid mutants.

@@

For Python, as with Java, many mutants we inspected for {\bf RQ5} involved deletion of a call or change in comparison operator, leading to a likely method call or access to a variable containing {\tt None}, and the generated finding reported just this problem.  Another example of a possibly engine-based mutation difference is that pyflakes did a much better job than the other tools of detecting cases where a mutant caused an exception to be unwisely ignored; these accounted for more than 10\% of differentially-relevant mutants; these may not all be faults, but are certainly worth inspecting.  We assume the issue is engine-based in that while pyflakes did by far the best at this (real) problem, there were instances where another tool found the same notional problem and it did not.  On the other hand, pyflakes, in the sample we examined, did worst at detecting mutants that removed a needed argument from a function call, another common pattern (Pylint performed best for this type of mutant).  Again, based on a qualitative examination of many mutants, we believe the results to primarily ($>$ 75\%) rest on realistic faults, which some tools correctly detected and others did not.
